Probabilistic Symbol Encoding for \\Convolutional Associative Memories

### 2017-08-01 papers/hrr-paper-neurotechnix/Example.tex

The symbolic scripting experiment is set up using a single-sensor Braitenberg
vehicle (see figures \ref{fig:experiment:agent1} and
\ref{fig:experiment:agent2}).
The sensor returns a list of objects in its field of view with their
corresponding view space positions in $[0,1]$ (far left to far right).
% note: it would be better if we probe with all returned objects and check the
% result, but IIRC this did not work well. maybe figure it out?
All returned objects are then compared to the agent's target object.
If the target is found, the holographic controller is probed with its view space
position. This position yields results even for
values that lie between the originally stored values due to the nature of scalar encoding and can consequently be
probed with the symbols $left\_wheel$ and $right\_wheel$ to extract the symbolic
actions $forwards$ or $backwards$.
If the target is not found or no action could be determined, the agent defaults
to probing the controller with the symbolic constant $no\_object$ which results in
a clockwise rotation when probing for wheel actions (see figure
\ref{fig:experiment:similar2}).

In the left column of figure \ref{fig:experiment}, the agent approaches the blue
target (figure \ref{fig:experiment:agent1}). Following figure {fig:experiment:pos1} in the beginning (tick $-50$), the
agent rotates right (default action) because the target is not visible.
At tick $-30$, the object enters the agent's visual field and continuously nears the fields center, which is reached at tick $-19$.
Then, with the sensory value of ${\sim}0.5$, the actions for both wheels switch
to $forwards$ (figure \ref{fig:experiment:similar1}) and the agent approaches
the target, constantly decreasing its distance
(figure \ref{fig:experiment:distance1}). In the right column of figure \ref{fig:experiment}, the target is not yet
in the visual field of the agent.
This is the result of the agent previously reaching an old target (at approximately tick $-13$) which lead
to the removal of all current objects and random placement of new ones,
including a new target.
As seen in figure {fig:experiment:pos2} since the new target is placed outside of the agent's field of view, the default
action of probing $no\_object$ is performed. The
clockwise rotation can also be seen in figure \ref{fig:experiment:similar2}). Regardless of how many times the agent reaches its target it will continue to find and pursue new ones solely with the help of its symbolic controller.

This experiment shows the effectiveness of our approach in a practical
application by encoding symbolic sensing, actuators and actions in a
controller.

@@

	\section{CONCLUSIONS}
	We have introduced a mechanism for intrinsic symbol encoding/decoding for Holographic Reduced Representations.
As a consequence, our approach eliminates the need for clean-up memories as they are described in literature.
Instead, results of probing operations directly give interpretable symbols, which can be decoded in constant computational time.
In other words, time needed for encoding/decoding is completely independent of the total number of symbols memorized.\\
	A concrete biophysical evidence and theoretical model of implementation of convolution in the brain is, unfortunately, still missing. Nevertheless, our work assumes convolution as the basic computational substrate of higher level cognition and identifies random permutation as computational requirement without which association of probabilistically encoded symbols is not possible.

>>>>

The symbolic scripting experiment is set up using a single-sensor Braitenberg
vehicle (see figures \ref{fig:experiment:agent1} and
\ref{fig:experiment:agent2}).
The sensor returns a list of objects in its field of view with their
corresponding view space positions in $[0,1]$ (far left to far right).
% note: it would be better if we probe with all returned objects and check the
% result, but IIRC this did not work well. maybe figure it out?
All returned objects are then compared to the agent's target object.
If the target is found, the holographic controller is probed with the targets
position. This position yields results even for
values that lie between the originally stored values due to the nature of scalar encoding and can consequently be
probed with the symbols $left\_wheel$ and $right\_wheel$ to extract the symbolic
actions $forwards$ or $backwards$.
If the target is not found or no action could be determined, the agent defaults
to probing the controller with the symbolic constant $no\_object$ which results in
a clockwise rotation when probing for wheel actions (see figure
\ref{fig:experiment:similar2}).

In the left column of figure \ref{fig:experiment}, the agent approaches the blue
target (figure \ref{fig:experiment:agent1}). Following figure \ref{fig:experiment:pos1} in the beginning (tick $-50$), the
agent rotates right (default action) because the target is not visible.
At tick $-30$, the object enters the agent's visual field and continuously nears the fields center, which is reached at tick $-19$.
Then, with the sensory value of ${\sim}0.5$, the actions for both wheels switch
to $forwards$ (figure \ref{fig:experiment:similar1}) and the agent approaches
the target, constantly decreasing its distance
(figure \ref{fig:experiment:distance1}). In the right column of figure \ref{fig:experiment}, the target is not yet
in the visual field of the agent.
This is the result of the agent previously reaching an old target (at approximately tick $-13$) which lead
to the removal of all current objects and random placement of new ones,
including a new target.
As seen in figure \ref{fig:experiment:pos2}, since the new target is placed outside of the agent's field of view, the default
action of probing $no\_object$ is performed. The
clockwise rotation can also be seen in figure \ref{fig:experiment:similar2}).

Regardless of how many times the agent reaches its target it will continue to find and pursue new ones solely with the help of its symbolic controller. This experiment shows the effectiveness of our approach in a practical
application by encoding symbolic sensing, actuators and actions in a
controller.

@@

	\section{CONCLUSIONS}
	We have introduced a mechanism for intrinsic symbol encoding/decoding for Holographic Reduced Representations.
As a consequence, our approach eliminates the need for clean-up memories as they are described in literature.
Instead, results of probing operations directly give interpretable symbols, which can be decoded in constant computational time.
In other words, time needed for encoding/decoding is completely independent of the total number of symbols memorized.\\
	While concrete biophysical evidence and a theoretical model of convolution in the brain is not yet available our work assumes convolution as the basic computational substrate of higher level cognition and identifies random permutation as a solution for a computational requirement without which association of probabilistically encoded symbols is not possible.
